\documentclass[letterpaper,10pt]{article}

\usepackage{latexsym}
\usepackage[empty]{fullpage}
\usepackage{titlesec}
\usepackage{marvosym}
\usepackage[usenames,dvipsnames]{color}
\usepackage{verbatim}
\usepackage{enumitem}
\usepackage{hyperref}
\hypersetup{colorlinks=true,urlcolor=cyan}
\usepackage{fancyhdr}
\usepackage[english]{babel}
\usepackage{tabularx}
\usepackage{multicol}
%\input{glyphtounicode}
\usepackage [autostyle, english = american]{csquotes}
\MakeOuterQuote{"}

%\usepackage[default]{sourcesanspro}
%\usepackage[T1]{fontenc}

\pagestyle{fancy}
\fancyhf{} 
\fancyfoot{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}


\addtolength{\oddsidemargin}{-0.5in}
\addtolength{\evensidemargin}{-0.5in}
\addtolength{\textwidth}{1in}
\addtolength{\topmargin}{-.5in}
\addtolength{\textheight}{1.0in}

\urlstyle{same}

\raggedbottom
\raggedright
\setlength{\tabcolsep}{0in}

\titleformat{\section}{
  \vspace{-4pt}\centering
}{}{0em}{}[\color{black}\titlerule\vspace{-5pt}]


\pdfgentounicode=1

\newcommand{\resumeItem}[1]{
  \item\small{
    {#1 \vspace{-2pt}}
  }
}

\newcommand{\resumeSubheading}[4]{
  \vspace{-2pt}\item
    \begin{tabular*}{0.97\textwidth}[t]{l@{\extracolsep{\fill}}r}
      \textbf{#1} & #2 \\
      \textit{\small#3} & \textit{\small #4} \\
    \end{tabular*}\vspace{-7pt}
}

\newcommand{\resumeSubSubheading}[2]{
    \item
    \begin{tabular*}{0.97\textwidth}{l@{\extracolsep{\fill}}r}
      \textit{\small#1} & \textit{\small #2} \\
    \end{tabular*}\vspace{-7pt}
}

\newcommand{\resumeProjectHeading}[2]{
    \item
    \begin{tabular*}{0.97\textwidth}{l@{\extracolsep{\fill}}r}
      \small#1 & #2 \\
    \end{tabular*}\vspace{-7pt}
}

\newcommand{\resumeSubItem}[1]{\resumeItem{#1}\vspace{-4pt}}

\renewcommand\labelitemii{$\vcenter{\hbox{\tiny$\bullet$}}$}

\newcommand{\resumeSubHeadingListStart}{\begin{itemize}[leftmargin=0.15in, label={}]}
\newcommand{\resumeSubHeadingListEnd}{\end{itemize}}
\newcommand{\resumeItemListStart}{\begin{itemize}}
\newcommand{\resumeItemListEnd}{\end{itemize}\vspace{-5pt}}

\begin{document}



\begin{center}
    {\LARGE Jacob Toner Gosselin} \\ \vspace{2pt}
    \begin{multicols}{2}
    \begin{flushleft}
    \href{{https://github.com/jaketg97/Jacob-Toner-Gosselin}}{Github}\\
    \href{{https://www.linkedin.com/in/jacobgosselin/}}{Linkedin}
    \end{flushleft}
    
    \begin{flushright}
    \href{{https://jaketg97.github.io/Jacob-Toner-Gosselin/}}{Personal site}\\
    \href{mailto:{jaketg97@gmail.com}}{Email}
    \end{flushright}
    \end{multicols}
\end{center}


%-----------EDUCATION-----------
\vspace{-2pt}
\section{Education}
  \resumeSubHeadingListStart
      \resumeSubheading
      {The University of Chicago}{Chicago, IL}
      {BA with Honors in Economics, BS in Mathematics, minor in Creative Writing}{June 2019}
      \resumeItemListStart
        \resumeItem{Graduated with general honors and departmental honors in economics; Dean’s List 2015-2019}
        \resumeItem{Honors Thesis: Insurer Competition in the Age of Provider Consolidation}
        \resumeItem{Relevant Coursework: Honors Econometrics, Probability Theory, Analysis in R$^n$ I-III, Stat. Methods and Models}
   	  \resumeItemListEnd

  \resumeSubHeadingListEnd


%-----------EXPERIENCE-----------
\section{Experience}
  \resumeSubHeadingListStart
    \resumeSubheading
      {Northwestern University}{Chicago, IL}
      {Research Analyst, Professor Seema Jayachandran}{July 2020-Present}
      \resumeItemListStart
        \resumeItem{Research Analyst for Professor Seema Jayachandran, working on research in applied microeconomics.}
        \resumeItem{Cleaned survey/administrative datasets and prepared them for analysis, conducted analysis using different statistical
and econometric tools (mostly using Stata and \textbf{\textsf{R}}), and prepared tables and figures with empirical results for publications and presentations (using \LaTeX).}
    \resumeItemListEnd

    \resumeSubheading
      {AmeriCorps}{Indianapolis, IN}
      {VISTA, IndyEast Promise Zone}{June 2019-July 2020}
      \resumeItemListStart
      	\resumeItem{"Volunteer in Service to America" (VISTA) with IndyEast Promise Zone, a federally designated "Promise Zone" on the east side of Indianapolis.}
        \resumeItem{Worked with Indiana University Public Policy Institute and state offices to generate accurate community data.}
        \resumeItem{Built capacity with community partners for afterschool programming, technical certification programming, etc...}
    \resumeItemListEnd
    
    \resumeSubheading
      {The University of Chicago}{Chicago, IL}
      {Research Assistant, Professor Neale Mahoney}{December 2017-August 2018}
      \resumeItemListStart
        \resumeItem{Research Assistant for Professor Neale Mahoney, working on research in health economics.}
        \resumeItem{Cleaned and analyzed data from the Healthcare Cost Report Information System in Stata for an ongoing research
project looking at the impact of monopsony in hospitals on lower wage workers.}
	\resumeItem{Performed literature reviews for presentation on the Affordable Care Act (ACA) at the Booth Business School.} 
    \resumeItemListEnd
    
     \resumeSubheading
      {Injustice Watch}{Chicago, IL}
      {Data Reporting Intern}{June 2018-June 2019}
      \resumeItemListStart
        \resumeItem{Published investigation of judicial misconduct and harsh sentencing in \href{https://chicago.suntimes.com/2018/8/17/18433578/cook-county-judge-seeking-retention-had-34-decisions-struck-down-in-6-years}{The Chicago Sun-Times}.}
        \resumeItem{Analyzed sentencing patterns of judges seeking re-election in \textbf{\textsf{R}} for voting guide published in \href{https://www.injusticewatch.org/interactives/judicial-guide/index.html}{The Chicago Sun-Times}.}
	\resumeItem{Mapped police stops using R and Tableau to analyze the effect of race and income on who “stop and frisk” impacts.} 
    \resumeItemListEnd
    
      \resumeSubheading
      {The Brookings Institution}{Washington, DC}
      {Economic Studies Intern}{June 2017-August 2017}
      \resumeItemListStart
        \resumeItem{Served as research assistant on three studies of rural healthcare under supervision of Dr. Alice Rivlin.}
        \resumeItem{Conducted statistical analysis on NSDUH data to determine impact of Medicaid enrollment on opioid users.}
	\resumeItem{Performed literature reviews to provide preliminary research for ongoing projects.} 
    \resumeItemListEnd

  \resumeSubHeadingListEnd
  
  %-----------EXPERIENCE-----------
\section{Leadership and Activities}
  \resumeSubHeadingListStart
    \resumeSubheading
      {The Institute of Politics}{The University of Chicago}
      {Civic Engagement Chair, Student Advisory Board}{June 2018-June 2019}
      \resumeItemListStart
        \resumeItem{Managed the interests of all student run organizations within the IOP, serving as a liaison between staff and students.}
        \resumeItem{Reviewed applications for new IOP student organizations, and helped coordinate funding.}
        \resumeItem{Represented students in weekly meetings with IOP staff, and quarterly meetings with IOP Director David Axelrod.}
    \resumeItemListEnd

    \resumeSubheading
      {Varsity XC and Track and Field}{The University of Chicago}
      {Athlete, Captain}{August 2016-June 2019}
      \resumeItemListStart
      	\resumeItem{Managed a 20+ hour weekly commitment with a full course load year-round.}
        \resumeItem{USTFCCCA Academic All-American Selection (2017), UAA Conference All-Academic Selection ('15, '16, '17, '18).}
    \resumeItemListEnd

  \resumeSubHeadingListEnd

%-----------PUBLICATIONS-----------
\section{Publications}
 \begin{itemize}[leftmargin=0.15in, label={}]
    \small{\item{
    \textbf{Insurer Competition in the Age of Provider
Consolidation: The relationship between hospital
and insurance competition in the ACA individual
market} - Honors Thesis, and updated as working paper on personal site {\\Jacob Toner Gosselin}\\
}}
 \end{itemize}

%-----------PROGRAMMING SKILLS-----------
\section{Technical Skills, Language Skills, and Interests}
 \begin{itemize}[leftmargin=0.15in, label={}]
    \small{\item{
     \textbf{Programming Languages}{: Experienced in Stata, \textbf{\textsf{R}}, \LaTeX, Python, Excel and Git for version control.} \\
     \textbf{Languages}{: Spanish (conversational)} \\
     \textbf{Volunteering}{: Former AmeriCorps VISTA; volunteer tutor with Chicago Peace Corps throughout college.} \\
     \textbf{Interests}{: Running (doing it and watching it); Basketball(just watching it, mostly NBA but some college); Writing.}
     
    }}
 \end{itemize}

\end{document}